Quality Assesment of OpenStreetMap Footpath Data for 3D City Recreation

### 2020-06-09 report/report.tex
Convert related works sections to subswctions
Move procedural methods section to Background, add todos
Move geodata qualification content from Implementation to Background

\section{Previous work in qualifying OpenStreetMap data}

@@

\section{Procedural models for city reproduction}

\section{Methods for geodata qualification}

\section{Level of detail}

@@

@@

\subsection{Assesment of dataset completeness}

\textbf{(Fan et al, 2014)} used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
The motivation for this is to eliminate semantic differences between both datasets.
A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
This makes a complete one-to-one object mapping impossible.
Rather, buildings have to be identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
Figure \ref{fig:building-match-types} shows examples of two buildings that were found in the OSM and SLU datasets as 1:1 and 1:n matches respectively.

Since it is impossible to find a one-to-one feature map for building footprints, using building area as an estimator for completeness is much better than i.e. using the number of buildings or other objects in each dataset.
This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.

@@

\subsection{Correspondence by building area}

\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
Figure \ref{fig:osm-slu-map} under Section 5 shows a typical comparison between the OSM and SLU datasets.
By eyesight it was determined that the displacement of features is small.

The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
\begin{center}
    $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
\end{center}

The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, \textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)} found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
Therefore it will here be assumed that two buildings are matching candidates if their relative area overlap is greater than 30\%.
If a one-to-many object matching is found, the compound perimeter of the footprints in the many-set will be used when calculating the shape accuracy and finding closest vertices between the polygons.

@@

The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
This project uses the turning function as it is defined by \textbf{(Fan et al, 2014)}.
See figure \ref{fig:turning-function} for a side-by-side comparison of a polygon and its turning function.
For a polygon with vertices ${v_1 ... v_n}$ and line segments ${e_1 ... e_n}$ It is defined as follows.
Fix a starting vertex $v_1$.
The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighbouring line segments $e_n$ and $e_1$.
For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as:
\begin{center}
    $\theta_{i, i-1} = \theta_{n,1} + \sum^{i}_{k=1} \theta_{k, k-1}$
\end{center}
The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon. The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under roation. It also measures only the normalized cumulative length, which does not change under scaling.
The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions:
\begin{center}
    $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
\end{center}
The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical. 
See figure \ref{fig:turning-function-compare} for a side-by-side comparison of two similar polygons and their turning functions.

@@

Procedural techniques are, as mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required. Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
\textbf{(Mũller et al, 2006)} present how to accomplish this using Shape Grammar: a method for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
For simple presentation in e.g. animation and videogames or even urban planning software, such procedurally generated cities could very well be enough.
As said however, the big problem is that these features will rarely match real life, such models will be sufficient as long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
(\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)})
and reducing noise
(\textbf{(Joon, Seo and Byung, 2011)}).

>>>>

\section{Related work}

\subsection{OpenStreetMap data qualification}

@@

\subsection{Procedural models for city reproduction}

Procedural techniques are, as mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required. Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
\textbf{(Mũller et al, 2006)} present how to accomplish this using Shape Grammar: a method for automatically and iteratively constructing 3D models based in a hierarchical set of rules.

By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.

%TODO: Mention a variety of procedural methods from different sources, start by reading Müller et al and look for more references

\subsection{Methods for geodata qualification}

\textbf{(Fan et al, 2014)} used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
The motivation for this is to eliminate semantic differences between both datasets.
A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
This makes a complete one-to-one object mapping impossible.
Since it is impossible to find a one-to-one feature map for building footprints, using building area as an estimator for completeness is much better than i.e. using the number of buildings or other objects in each dataset.

\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
\begin{center}
    $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
\end{center}

The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other,
\textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)}
found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.

%TODO: DP, describe the algorithm with reference to DP and Fan
Douglas-Peucker algorithm for polygon reduction. The full procedure for matching points will be presented in section ***.

%TODO: Turning function for shape accuracy, find source.
The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.

%TODO: Read Fan et al and look for references

\subsection{Level of detail}

@@

%TODO: Summarize both sources by Biljecki

@@

\subsection{Dataset completeness and correspondence}

In the same way as the study by \textbf{(Fan et al, 2014)}, this project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
The motivation for this is to eliminate semantic differences between both datasets.
Buildings were identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
Figure \ref{fig:building-match-types} shows examples of two buildings that were found in the OSM and SLU datasets as 1:1 and 1:n matches respectively.

@@

I accordance with \textbf{Fan et al, 2014}, 
In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.

Figure \ref{fig:osm-slu-map} under Section 5 shows a typical comparison between the OSM and SLU datasets.

\textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)}, 30% area overlap
Therefore it will here be assumed that two buildings are matching candidates if their relative area overlap is greater than 30\%.
If a one-to-many object matching is found, the compound perimeter of the footprints in the many-set will be used when calculating the shape accuracy and finding closest vertices between the polygons.

@@

The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
This project uses the turning function as it is defined by \textbf{(Fan et al, 2014)}.
See figure \ref{fig:turning-function} for a side-by-side comparison of a polygon and its turning function.
For a polygon with vertices ${v_1 ... v_n}$ and line segments ${e_1 ... e_n}$ It is defined as follows.
Fix a starting vertex $v_1$.
The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighbouring line segments $e_n$ and $e_1$.
For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as:
\begin{center}
    $\theta_{i, i-1} = \theta_{n,1} + \sum^{i}_{k=1} \theta_{k, k-1}$
\end{center}
The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon. The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under roation. It also measures only the normalized cumulative length, which does not change under scaling.
The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions:
\begin{center}
    $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
\end{center}
The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical. 
See figure \ref{fig:turning-function-compare} for a side-by-side comparison of two similar polygons and their turning functions.

@@

For simple presentation in e.g. animation and videogames or even urban planning software, such procedurally generated cities could very well be enough.
As said however, the big problem is that these features will rarely match real life, such models will be sufficient as long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
(\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)})
and reducing noise
(\textbf{(Joon, Seo and Byung, 2011)}).
